---
title: "EFleming COMM 310L Final"
author: "Emerson Fleming"
date: "5/9/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(readxl)
library(ggcorrplot)
library(car)
library(ggpubr)
```

```{r}
library(readr)
TV_Data <- read_csv("/Applications/COMMFINAL/EFleming Dataset (new) - EFleming Dataset (new).csv")
```

**Filtering for people that selected Bachelor as at least one of their favorite shows**

```{r}
Bachelor_as_fav <- filter(TV_Data, TV_most_1_TEXT == "The Bachelor" | TV_most_2_TEXT == "The Bachelor" | TV_most_3_TEXT == "The Bachelor" | TV_most_4_TEXT == "The Bachelor" | TV_most_5_TEXT == "The Bachelor")
```

**Filtering for people that DID NOT select Bachelor as at least one of their favorite shows**

```{r}
Bachelor_not_as_fav <- filter(TV_Data, TV_most_1_TEXT != "The Bachelor" & TV_most_2_TEXT != "The Bachelor" & TV_most_3_TEXT != "The Bachelor" & TV_most_4_TEXT != "The Bachelor" & TV_most_5_TEXT != "The Bachelor")
```

**Creation of variable for correlation matrix for people that selected Bachelor as one of their favorite shows**

```{r}
TV_Data_Cor_as_fav <- select(Bachelor_as_fav, "Bachelor_Favoritism", "Age", "Sex", "TV", "Spir", "Tvsoap", "Tvdrama", "Tvsitcom", "Porn3", "Single", "TV_sex1", "TV_sex2", "TV_sex3")
TV_Data_Cor_as_fav <- TV_Data_Cor_as_fav %>% drop_na()
```

**Creation of variable for correlation matrix of entire sample (people that did and did not select Bachelor)**

```{r}
TV_Data_Cor_Whole_Sample <- select(TV_Data, "Bachelor_Favoritism", "Age", "Sex", "TV", "Spir", "Tvsoap", "Tvdrama", "Tvsitcom", "Porn3", "Single", "TV_sex1", "TV_sex2", "TV_sex3")
TV_Data_Cor_Whole_Sample <- TV_Data_Cor_Whole_Sample %>% drop_na()
```

**Creation of variable for correlation matrix for people that did and did not select Bachelor as one of their favorite shows**

```{r}
TV_Data_Cor_not_as_fav <- select(Bachelor_not_as_fav, "Bachelor_Favoritism", "Age", "Sex", "TV", "Spir", "Tvsoap", "Tvdrama", "Tvsitcom", "Porn3", "Single", "TV_sex1", "TV_sex2", "TV_sex3")
TV_Data_Cor_not_as_fav <- TV_Data_Cor_not_as_fav %>% drop_na()
```

### Correlation Matrix for those that include The Bachelor as their favorite
---->This correlation matrix includes a new "Bachelor_Favoritism" ordinal variable in which the preference for The Bachelor is used on a scale of 0-5 (based on where people rank The Bachelor).

```{r}
Cor_Data_As_Fav <- cor(TV_Data_Cor_as_fav) 
ggcorrplot(Cor_Data_As_Fav, lab = TRUE, lab_size = 2)
```

### Correlation Matrix for Whole Sample
---->This correlation matrix includes a new "Bachelor_Favoritism" ordinal variable in which the preference for The Bachelor is used on a scale of 0-5 (based on where people rank The Bachelor).

```{r}
Cor_Data_Whole_Sample <- cor(TV_Data_Cor_Whole_Sample) 
ggcorrplot(Cor_Data_Whole_Sample, lab = TRUE, lab_size = 2) 
```

**We created a new variable here (a dummy variable for Bachelor Favoritism)**
```{r}
TV_Data$Bachelor = ifelse(TV_Data$Bachelor_Favoritism>0,1,0)
Bachelor_glm_model <- glm(Bachelor~Age + Sex + TV + Spir + Tvdrama + Tvsitcom + Porn3 + Single, data = TV_Data, family = "binomial")
summary(Bachelor_glm_model)
```

```{r}
Bachelor_model <- lm(Bachelor_Favoritism~Age + Sex + TV + Spir + Tvdrama + Tvsitcom + Porn3 + Single, data = TV_Data)
summary(Bachelor_model)
```

$$
Bachelor_model <- lm(Bachelor_Favoritism~Age + Sex + TV + Spir + Tvdrama + Tvsitcom + Porn3 + Single, data = TV_Data)
$$


```{r}
Bachelor_model_2 <- lm(Bachelor_Favoritism~Spir, data = TV_Data[TV_Data$Bachelor==1,])
```


**Function to find and get rid of multicollinearity**
```{r}
car::vif(Bachelor_model)
```

*Deleted TV_sex1, TV_sex2, and TV_sex3 as these are likely correlated with porn3 and these are interfering with our p-values when we run a regression.

*We also can drop TV Soap as this value is very similar to TV drama. Therefore, there isn't any real need to include this in the regression.

### Still we are not super successful with creating statistical significance

```{r}
TV_Data$Spirtual_new = ifelse(TV_Data$Spir>50,1,0)
Bachelor_as_fav$Spirtual_new = ifelse(Bachelor_as_fav$Spir>50,1,0)
Spir_1 = TV_Data[TV_Data$Spirtual_new==1,]
Spir_0 = TV_Data[TV_Data$Spirtual_new==0,]
Spir_1_Cor <- cor(na.omit(Spir_1[,c("Bachelor_Favoritism", "Age", "Sex", "TV", "Tvsoap", "Tvdrama", "Tvsitcom", "Porn3", "Single", "TV_sex1", "TV_sex2", "TV_sex3", "Spir")]))
Spir_0_Cor <- cor(na.omit(Spir_0[,c("Bachelor_Favoritism", "Age", "Sex", "TV", "Tvsoap", "Tvdrama", "Tvsitcom", "Porn3", "Single", "TV_sex1", "TV_sex2", "TV_sex3", "Spir")]))
```

When you control for spirituality, other variables are not highly correlated to Bachelor_Favoritism. Here create 2 groups of people. One for people with spirituality above median (50) and one below 50. Once you control for spirituality, the correlations go away. Correlation is being driven not by Bachelor Favoritism and each variable by themselves but spirituality.

```{r}
ggcorrplot(Spir_1_Cor, lab = TRUE, lab_size = 2) 
```

```{r}
ggcorrplot(Spir_0_Cor, lab = TRUE, lab_size = 2) 
```


```{r}
ggplot(Bachelor_as_fav,
  aes(x = Spir, y = Bachelor_Favoritism)) +
  coord_cartesian(xlim = c(1, 100), ylim = c(0, 5)) +
  geom_point(size = 0.5) +
  geom_line(colour = "red") +
  geom_smooth(colour = "black") +
  labs(title = "Sprituality vs. Bachelor Favoritism",
       x = "Spirituality",
       y = "Bachelor Favoritism")+
  theme_classic()
```
Here, we simply run an exploratory plot based only on those that selected the Bachelor as one of their top 5 favorite shows (as denoted by Bachelor Favoritism). As we can see based only on the original data, we can see that the least spiritual people seem to rank The Bachelor as their favorite show. However, those that still rank the Bachelor as their 2nd or 3rd favorite show exhibit higher levels of spirituality as well.

It is important to note the lack of observations and their impact on this result.

### Model Representation for Spirituality vs. Bachelor Favoritism

```{r}
ggplot(Bachelor_model_2,
  aes(x = Spir, y = Bachelor_Favoritism)) +
  geom_point() +
  geom_abline(slope = coef(Bachelor_model_2)[["Spir"]], 
              intercept = coef(Bachelor_model_2)[["(Intercept)"]], color = "red") +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. Spirituality",
       x = "Spirituality",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))
```

For our model we have a slight negative relationship between Bachelor Favoritism and Spirituality.

The important thing is that we are demonstrating that while we do have high correlations, they are not actually important when we visualize the model. We had the highest statistical significance for the p-value between Bachelor Favoritism and Spirituality. However, there small statistical significance when graphically portrayed.

We have chosen Spirituality to represent as this is the only variable in the model that exhibits statistical significance (a p-value smaller than 0.05).

**Models for Bachelor Favoritism vs. TV Drama plot**
```{r}
Bachelor_model_3 <- lm(Bachelor_Favoritism~Tvdrama, data = TV_Data)
Bachelor_model_4 <- lm(Bachelor_Favoritism~Tvdrama, data = Spir_0,)
Bachelor_model_5 <- lm(Bachelor_Favoritism~Tvdrama, data = Spir_1)
```

### Representation of Bachelor Favoritism vs. TV Drama
```{r}
ggplot() +
  geom_point(data = Bachelor_model_3, aes(x = Tvdrama, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_3, slope = coef(Bachelor_model_3)[["Tvdrama"]], 
              intercept = coef(Bachelor_model_3)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_4, aes(x = Tvdrama, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_4, slope = coef(Bachelor_model_4)[["Tvdrama"]], 
              intercept = coef(Bachelor_model_3)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_5, aes(x = Tvdrama, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_5, slope = coef(Bachelor_model_5)[["Tvdrama"]], 
              intercept = coef(Bachelor_model_3)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 15), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. TV Drama",
       x = "TV Drama",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))

q1 <- ggplot() +
  geom_point(data = Bachelor_model_3, aes(x = Tvdrama, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_3, slope = coef(Bachelor_model_3)[["Tvdrama"]], 
              intercept = coef(Bachelor_model_3)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_4, aes(x = Tvdrama, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_4, slope = coef(Bachelor_model_4)[["Tvdrama"]], 
              intercept = coef(Bachelor_model_3)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_5, aes(x = Tvdrama, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_5, slope = coef(Bachelor_model_5)[["Tvdrama"]], 
              intercept = coef(Bachelor_model_3)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 15), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. TV Drama",
       x = "TV Drama",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))
```
For the following models, the dataset was split into three groups. The blue line represents the entire sample with regards to those who selected The Bachelor as one of their favorite shows and those who did not.The green line represents those with 50% or higher levels of spirituality. The red line represents represents those with 50% or lower levels of spirituality. 

The "TV Soap" and "TV Drama" variables had very similar levels of correlation. Therefore, only TV Drama was included in the model to ensure the highest levels of statistical significance. 

As we can observe, individuals with higher levels of spirituality that watched the most TV Drama tend have higher levels of of Bachelor Favoritism compared to other groups. 

**Models for Bachelor Favoritism vs. TV Sitcom plot**
```{r}
Bachelor_model_6 <- lm(Bachelor_Favoritism~Tvsitcom, data = TV_Data)
Bachelor_model_7 <- lm(Bachelor_Favoritism~Tvsitcom, data = Spir_0)
Bachelor_model_8 <- lm(Bachelor_Favoritism~Tvsitcom, data = Spir_1)
```

### Representation of Bachelor Favoritism vs. TV Drama plot
```{r}
ggplot() +
  geom_point(data = Bachelor_model_6, aes(x = Tvsitcom, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_6, slope = coef(Bachelor_model_6)[["Tvsitcom"]], 
              intercept = coef(Bachelor_model_6)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_7, aes(x = Tvsitcom, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_7, slope = coef(Bachelor_model_7)[["Tvsitcom"]], 
              intercept = coef(Bachelor_model_7)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_8, aes(x = Tvsitcom, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_8, slope = coef(Bachelor_model_8)[["Tvsitcom"]], 
              intercept = coef(Bachelor_model_8)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 15), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. TV Sitcom",
       x = "TV Sitcom",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))

q2 <- ggplot() +
  geom_point(data = Bachelor_model_6, aes(x = Tvsitcom, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_6, slope = coef(Bachelor_model_6)[["Tvsitcom"]], 
              intercept = coef(Bachelor_model_6)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_7, aes(x = Tvsitcom, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_7, slope = coef(Bachelor_model_7)[["Tvsitcom"]], 
              intercept = coef(Bachelor_model_7)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_8, aes(x = Tvsitcom, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_8, slope = coef(Bachelor_model_8)[["Tvsitcom"]], 
              intercept = coef(Bachelor_model_8)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 15), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. TV Sitcom",
       x = "TV Sitcom",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))
```
All groups with regards to Bachelor favoritism exhibit a small and negative relationship between Bachelor Favoritism and TV Sitcom watched. 

**Models for Bachelor Favoritism vs. TV Soap Opera plot** 
```{r}
Bachelor_model_9 <- lm(Bachelor_Favoritism~Tvsoap, data = TV_Data)
Bachelor_model_10 <- lm(Bachelor_Favoritism~Tvsoap, data = Spir_0)
Bachelor_model_11 <- lm(Bachelor_Favoritism~Tvsoap, data = Spir_1)
```

### Representation of Bachelor Favoritism vs. TV Soap Opera plot
```{r}
ggplot() +
  geom_point(data = Bachelor_model_9, aes(x = Tvsoap, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_9, slope = coef(Bachelor_model_9)[["Tvsoap"]], 
              intercept = coef(Bachelor_model_9)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_10, aes(x = Tvsoap, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_10, slope = coef(Bachelor_model_10)[["Tvsoap"]], 
              intercept = coef(Bachelor_model_10)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_11, aes(x = Tvsoap, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_11, slope = coef(Bachelor_model_11)[["Tvsoap"]], 
              intercept = coef(Bachelor_model_11)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. TV Soap",
       x = "TV Soap",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))

q3 <- ggplot() +
  geom_point(data = Bachelor_model_9, aes(x = Tvsoap, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_9, slope = coef(Bachelor_model_9)[["Tvsoap"]], 
              intercept = coef(Bachelor_model_9)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_10, aes(x = Tvsoap, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_10, slope = coef(Bachelor_model_10)[["Tvsoap"]], 
              intercept = coef(Bachelor_model_10)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_11, aes(x = Tvsoap, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_11, slope = coef(Bachelor_model_11)[["Tvsoap"]], 
              intercept = coef(Bachelor_model_11)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. TV Soap",
       x = "TV Soap",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))
```

**Models for Bachelor Favoritism vs. Single plot**
```{r}
Bachelor_model_12 <- lm(Bachelor_Favoritism~Single, data = TV_Data)
Bachelor_model_13 <- lm(Bachelor_Favoritism~Single, data = Spir_0)
Bachelor_model_14 <- lm(Bachelor_Favoritism~Single, data = Spir_1)
```

### Bachelor Favoritism vs. Single plot 
```{r}
ggplot() +
  geom_point(data = Bachelor_model_12, aes(x = Single, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_12, slope = coef(Bachelor_model_12)[["Single"]], 
              intercept = coef(Bachelor_model_12)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_13, aes(x = Single, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_13, slope = coef(Bachelor_model_13)[["Single"]], 
              intercept = coef(Bachelor_model_13)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_14, aes(x = Single, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_14, slope = coef(Bachelor_model_14)[["Single"]], 
              intercept = coef(Bachelor_model_14)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. Single",
       x = "Single",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))

q4 <- ggplot() +
  geom_point(data = Bachelor_model_12, aes(x = Single, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_12, slope = coef(Bachelor_model_12)[["Single"]], 
              intercept = coef(Bachelor_model_12)[["(Intercept)"]], color = "red") +
  geom_point(data = Bachelor_model_13, aes(x = Single, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_13, slope = coef(Bachelor_model_13)[["Single"]], 
              intercept = coef(Bachelor_model_13)[["(Intercept)"]], color = "blue") +
  geom_point(data = Bachelor_model_14, aes(x = Single, y = Bachelor_Favoritism)) +
  geom_abline(data = Bachelor_model_14, slope = coef(Bachelor_model_14)[["Single"]], 
              intercept = coef(Bachelor_model_14)[["(Intercept)"]], color = "green") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 5)) +
labs(caption = "Source: UMass Amherst",
        title = "Bachelor Favoritism vs. Single",
       x = "Single",
       y = "Bachelor Favoritism") +
theme_minimal() +
theme(plot.title = element_text(size=15))
```

```{r}
ggarrange(q1, q2, nrow = 1)
```

```{r}
ggarrange(q3, q4, nrow = 1)

```